---
title: "Artificial Lift Selection Tool v0.1.0"
resource_files:
- inst/extdata/Artificial Lift Matrix for Commercial Production Systems (CPS)14Aug062.xlsx
runtime: shiny
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    source_code: embed
---

```{r setup, include=FALSE}
library(flexdashboard)
library(DT)
```


```{r rows.print=25}
library(readxl)
library(writexl)
library(dplyr)
library(tidyr)
library(ggplot2)


xls_file <- "Artificial Lift Matrix for Commercial Production Systems (CPS)14Aug062.xlsx"

al_matrix <- read_excel(path = file.path('inst/extdata', xls_file), 
                        skip = 4, range = "C5:P63") %>% 
    rename(design_aspect = "Design Aspect")


al_matrix_tidy <- al_matrix %>% 
    gather(key = 'al_method', value = "ranking", Jet:'Chamber Lift') %>% 
    mutate(ranking = as.integer(ranking)) %>% 
    mutate(ranking = ifelse(is.na(ranking), 0, ranking)) %>% 
    mutate(relevance = case_when(
        .$al_method %in% c("ESP", "Gas Lift", "Sucker Rod Pump", "PCP") ~ "high",
        .$al_method %in% c("Jet", "ESP-CP", "Sucker Rod Pump", "PCP", "HSP") ~ "low",
        .$al_method %in% c("Metal PCP (Hi Temp)", "HGP", 
                           "Twin Screw (ESTSP)", "Twin Screw (TDTSP)", "Chamber Lift") ~ "specialty"
    )) %>% 
    mutate(relevance = as.factor(relevance)) %>% 
    mutate(ranking = as.factor(ranking))

# write_xlsx(al_matrix_tidy, path = file.path('inst/extdata', "al_matrix_tidy.xlsx"))
al_matrix_tidy_orig <- al_matrix_tidy
```

# Sidebar {.sidebar data-width="350"}

```{r}
selectInput("design_aspect", "Design Aspect of the AL", 
            choices = unique(al_matrix_tidy$design_aspect))

# renderUI({
# selectInput("value", "Value of the Design Aspect to evaluate",
#             choices = rownames(al_matrix_tidy)[al_matrix_tidy$Value %in% input$design_aspect])
# })

renderUI({
    selectInput("value", "Value of the Design Aspect to evaluate",
                choices = al_matrix_tidy$Value[al_matrix_tidy$design_aspect == levels(al_matrix_tidy$design_aspect)[1]])
})

br()

observe({
    aspect <- input$design_aspect
    updateSelectInput(session, "value", 
                      choices = al_matrix_tidy$Value[al_matrix_tidy$design_aspect == aspect])
})    
```


**Info**

This is a selector of Artificial Lift methods by various operational Design Aspects. It is not the intention of being a rigorous selection tool since the criteria values are limited in number. See for instance, `Casing Size` with only two values of **7"** and **9.625" and larger**, or the `Relative Intervention Frequency` with only one value.

The selection tool may grow over time with feedback from the community of petroleum engineers that desire to bring this application to a new level of completeness.

**What is Relevance**

The meaning of Relevance here is the *popularity* or how much the Artificial Method is used in the industry where `high` should convey the understanding that the AL method is highly used, while `specialty` refers to niches areas of deployment.

**Ranking**
A higher value of the rankin means that the better the AL method will perform. The ranking number go from `1` (lower performance) through `3` (highest).



# Artificial Lift Matrix

## 

### AL Design Aspects {.value-box}

```{r}
renderValueBox({
    valueBox(length(unique(al_matrix_tidy$design_aspect)))
})
```

### AL Methods {.value-box}

```{r}
renderValueBox({
    valueBox(length(unique(al_matrix_tidy$al_method)))
})
```

### Evaluation Criteria  {.value-box}

```{r}
renderValueBox({
    valueBox(length(unique(al_matrix_tidy$Value)))
})
```

## Column

### Plot by Design Aspect

```{r warning=FALSE, message=FALSE, error=FALSE}

al_matrix_tidy <- al_matrix_tidy_orig

nameorder <- make.unique(al_matrix_tidy$al_method[order(al_matrix_tidy$relevance, al_matrix_tidy$ranking)])
    
al_matrix_tidy$al_method <- factor(al_matrix_tidy$al_method, levels=nameorder)

dataset <- reactive({
    subset(al_matrix_tidy, 
         design_aspect == input$design_aspect & Value == input$value )    
})


shiny::renderPlot({
    p <- ggplot(dataset(), aes(x = ranking, y = al_method)) 
    p <- p + geom_segment(aes(yend = al_method), xend = 0, color = "grey50") 
    p <- p + geom_point(size = 3, aes(colour = relevance))
    p <- p + scale_colour_brewer(palette="Set1", limits=c("specialty","low", "high")) 
    p <- p + theme_bw()
    p <- p + theme(panel.grid.major.y = element_blank())  # No horizontal grid lines
    p <- p + ggtitle(paste0("by ", input$design_aspect, " when " , input$value))
    # facets <- paste(dataset()$relevance, '~ .')
    # if (facets != '. ~ .')
    p <- p + facet_grid(relevance ~ ., scales="free_y", space = "free_y")
    print(p)
})
```


# Results on selection

##  {data-orientation=rows}

### Table by Design Aspects

```{r}
renderTable({
    subset(al_matrix_tidy, 
             design_aspect == input$design_aspect & Value == input$value )
})
```

##  {data-orientation=rows}

### Dynamic table

```{r}
renderDataTable({
    DT::datatable(subset(al_matrix_tidy, 
                 design_aspect == input$design_aspect & Value == input$value ),
        filter = "top",
        options = list(pageLength = 20, scrollY = "200px", paging = FALSE)
    )
})    
```
